---
title: "Interesting Forecasting Applications with R"
author: "Emirhan Esen - IE360 - Spring 2021"
---
```{r,echo=FALSE,results="hide",warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r,echo=FALSE,eval=FALSE,warning=FALSE}
rmarkdown::render("/home/baydogan/Courses/IE360/Guidelines/example_homework_0.Rmd",output_format="pdf_document")
rmarkdown::render("/home/baydogan/Courses/IE360/Guidelines/example_homework_0.Rmd",output_format="html_document")
```

1. [Time series forecasting using ARIMA models](https://blogs.oracle.com/datascience/introduction-to-forecasting-with-arima-in-r)

    This example is important because it introduces the forecasting methodology using ARIMA models in R. The steps are ...

2.[Using R for Time Series Analysis](https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html)

    An application for Time Series Analysis by using R 
    
3.[A Complete Tutorial on Time Series Modeling in R](https://www.analyticsvidhya.com/blog/2015/12/complete-tutorial-time-series-modeling/)

    There are basic for Time Series Modeling in R  and basic ARMA examples.
    
4.[Time Series Forecasting](http://r-statistics.co/Time-Series-Forecasting-With-R.html)    

    This is a follow-up to the introduction to time series analysis, but focused more on forecasting rather than analysis.
    
5.[Stock Price Forecasting Using Time Series Analysis](https://rpubs.com/kapage/523169)

    This is time series forecasting models applied to a stock close price. The forecasting method try to predict the next 30 days  close from today.